---
title: "Project Proposal"
output:
  html_document: 
    theme: readable
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The names of the students who will be contributing to the group project

Our group includes:Kiranmayee Nimashakavi, Gowri Shankar
Ramanan, Stefani Moore, and Yun Ying (Lomis) Chen


### A tentative title for the project.

Predicting King County house prices with multiple linear regression.

### Description of the data file (what they contain including number of variables and number of records).

The data set contains a total of 21614 records of housing prices of
Seattle King County, each with 21 columns/features/variables. Some
example of the variables include price, bedrooms, bathrooms, sqft_living, sqft_lot, year of build, and etc


```{r}

train_df <- read.csv("train.csv")
str(train_df)
```

###  Background information on the data sets, including specific citation of their source(so that I can also access it).

The dataset contains house sale prices for King County, which is a
county as part of seattle. The dataset includes homes sold between May 2014 and May 2015. Reference here: https://www.kaggle.com/harlfoxem/housesalesprediction


### Statement of personal interest

Statement of personal interest: A number of us have the aspiration to
learn more about applying learnings from this course to the chosen housing data set, including but not limited to data prep/cleaning, data exploration,data modeling, model optimization including using
forward/backwards/stepwise AIC/BIC, noise removals, as well as
classification and etc


### Evidence that data can be loaded into R

```{r}
head(train_df)

```

